---
title: "anti-trans-exercises"
author: "Group 00"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
bills.raw <- read.csv("/Users/pattihamerski/Downloads/anti-trans-bills-clean.csv")
```

```{r}
bills <- bills.raw %>%
  mutate(year = as.integer(substr(Date, 1, 4)),
         month = as.integer(substr(Date, 6, 7)),
         day = as.integer(substr(Date, 9, 10)),
         year.month = year + month/12,
         passed = ifelse(Status %in% c("Passed", "Signed/Enacted"), 1, 0)) %>% 
  filter(year >= 2017)

bills
```

How to explore data visually

Variables: 

Column             | Meaning
-------------------|----------------------------------------
`Date`             | date
`State`            | state in US
`Bill`             | identifier for chamber and bill number
`Bill.Description` | description of bill
`Bill.Type`        | type of anti-trans legislation
`Status`           | outcome of bill
`Bill.Link`        | URL to bill

Number of bills from each state (table)

```{r}
state.counts <- bills %>%
  group_by(State) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))

state.counts
```

Bills per month by type (line plot)


```{r}
type.counts <- bills %>%
  group_by(Bill.Type, year.month) %>% 
  summarize(count=n()) %>% 
  arrange(year.month)

type.counts
```

```{r}
bills %>%
  group_by(Bill.Type, year) %>% 
  summarize(count=n()) %>% 
  arrange(year) %>% 
  ggplot(aes(x = year, y = count)) +
  geom_line(aes(color = Bill.Type))
```



```{r}
bills %>%
  ggplot(aes(x = year)) + 
  geom_bar(aes(fill = Bill.Type), position = "dodge")
```



Number of bills of each type






Bills of each type and outcome of types

```{r}
ordered.status <- c("Introduced", "Crossed Over", "Passed",
                    "Vetoed", "Signed/Enacted", "Dead/Failed")

bills$Status <- ordered(bills$Status, levels = ordered.status)
```


```{r}
colors.status = c("blue", "cyan", "pink", "red", "magenta", "purple")

bills %>%
  ggplot(aes(x = Bill.Type)) + 
  geom_bar(aes(fill = Status), position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_manual(values = colors.status)
```



total num of bills proposed each year
total num of bills passed each year

```{r}
bills %>%
  group_by(year.month) %>% 
  summarize(count=n(), num.passed=sum(passed)) %>% 
  ggplot(aes(x = year.month, y = count)) +
  geom_point(aes(y=count), color="blue") +
  geom_point(aes(y=num.passed), color="red")
```







types of bills proposed in 2023 compared to previous years
types passed more often in 2023
which states introduced and/or passed most anti-trans bills






